---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# pharmacocaen

<!-- badges: start -->
<!-- badges: end -->

The goal of pharmacocaen is to provide tools for worldwide pharmacovigilance analysis.

## Installation

You can install the development version of pharmacocaen from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("cdolladille/pharmacocaen")
```

## Vignettes

There is a detailed vignette to explain simple data management [here](https://github.com/cdolladille/pharmacocaen/tree/master/vignettes/)

## Example

Say you want to create columns of drug and adr, then perform a univariate disproportionality analysis.
You may want to use the `add_drug`, `add_adr`, and `compute_or_abcd` functions.

```{r example, warning=FALSE, message=FALSE}
library(pharmacocaen)

demo <-
  demo_ %>%
  add_drug(
    d_code = ex_$d_drecno,
    drug_data = drug_
  ) %>%
  add_adr(
    a_code = ex_$a_llt,
    adr_data = adr_
  )

demo %>%
  compute_or_abcd(
    y = "colitis",
    x = "nivolumab"
  )
```

<!-- Footnote for myself

You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this. You could also use GitHub Actions to re-render `README.Rmd` every time you push. An example workflow can be found here: <https://github.com/r-lib/actions/tree/v1/examples>. -->
