---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# vigicaen

<!-- badges: start -->
[![R-CMD-check](https://github.com/pharmacologie-caen/vigicaen/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/pharmacologie-caen/vigicaen/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of vigicaen is to provide tools for worldwide pharmacovigilance analysis.

## Installation

You can install the development version of vigicaen from [GitHub](https://github.com/) with:

> `devtools::install_github("pharmacologie-caen/vigicaen")`

### Solution 2

Look at the **Releases** panel on the right of this Github page. 
You may find "vX.XX.X (Latest)".

Click on the version you want to install.

Download source code as a tar.gz file.

Now go to RStudio, click on "Tools", "Install Packages...", select 
"Package Archive file" and locate the tar.gz file on your computer.


## Vignettes

//WIP

## Example

Say you want to create columns of drug and adr, then perform a univariate disproportionality analysis.
You may want to use the `add_drug`, `add_adr`, and `compute_or_abcd` functions.

```{r example, warning=FALSE, message=FALSE}
library(vigicaen)

demo <-
  demo_ |> 
  add_drug(
    d_code = ex_$d_drecno,
    drug_data = drug_
  ) |> 
  add_adr(
    a_code = ex_$a_llt,
    adr_data = adr_
  )

demo |> 
  compute_or_abcd(
    y = "a_colitis",
    x = "nivolumab"
  )
```

<!-- Footnote for myself

You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this. You could also use GitHub Actions to re-render `README.Rmd` every time you push. An example workflow can be found here: <https://github.com/r-lib/actions/tree/v1/examples>. -->
