---
title: "build_tables"
output: 
  rmarkdown::html_vignette:
    keep_md: true
    toc: true
vignette: >
  %\VignetteIndexEntry{build_tables}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(pharmacocaen)
```

# Purpose

Have you subscribed to VigiBase Extract Case Level, WHODrug and MedDRA?

Congratulations! You've made it half way to explore VigiBase.

Now, you need to process those large text/ascii files into R readable files. This is the purpose of the `tb_*` functions (Table Builders).

## Overview

| tb\_\* function | Usage                                             |
|-----------------|---------------------------------------------------|
| tb_main         | Main tables (demo, drugs, adverse drug reactions) |
| tb_who          | WHODrug tables                                    |
| tb_sub          | VigiBase subsidiary tables                        |
| tb_meddra       | MedDRA tables                                     |

## Basic usage

You just need to provide a path to these files on your computer.

```{r basic, eval = FALSE}

tb_main(path_main, path_sub)
tb_who(path_who)
tb_sub(path_sub)
tb_meddra(path_meddra)

```

And this is a step you only have to perform ONCE per database version.

New parquet files will be created in the same directories as provided with `path_*`.

> The suspectedduplicates table is delivered with subsidiary files, but the `tb_main` function will output it in the main directory, under a slightly modified name.

## Outputs

The `tb_*` functions create [parquet](https://arrow.apache.org/docs/r/index.html) files. Parquet is an open-source format, supported by Arrow.

Not all tables from each source are created during the process.

-   Main: All tables + suspdup (from sub).

-   WHODrug: All tables - mp + mp_short.

-   Subsdiary files: All tables.

-   Meddra: meddra_hierarchy, smq_list, and smq_content.

mp_short is a short version of mp table, which has been tolower-ed and trim-ed on drug name, to be more handy. A few columns also have been removed.

# Loading tables into R

You should use the arrow package built-in readers to load the tables.

```{r load_tables, eval = FALSE}
library(arrow)

demo <- read_parquet(path_main, "demo.parquet")
```

> Parquet files will be loaded OUT of memory, by default. To change this behavior, remember to `dplyr::collect()` the tables

You can `left_join` smq_content to smq_list if you work only with NON-algorithmic SMQs.

```{r joining_smq_tables, eval = FALSE}
smq_list    <- read_parquet(path_meddra, "smq_list.parquet")
smq_content <- read_parquet(path_meddra, "smq_content.parquet")

smq_list_content <-
  smq_list |>
  # Only works for NON-algorithmic SMQs
  left_join(smq_content, by = "smq_code") |>
  data.table
```
