---
title: "generic_main"
output: 
  rmarkdown::html_vignette:
    keep_md: true
vignette: >
  %\VignetteIndexEntry{generic_main}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}

---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  include = TRUE,
  echo = TRUE,
  warning = FALSE, 
  message = FALSE
)
```

WELCOME to the generic process to perform a disproportionality analysis
using VigiBase ECL!

Please follow the steps in this script.

## Libraries

```{r libraries}
library(pharmacocaen) # requires data.table, dplyr, fst, rlang, tidyverse
library(rlang)
library(here) # if you like the here syntax for file paths
```

## Imports

You may create a separate script with paths, that you would call at the beginning
of your main script.

Set paths to your Vigibase ECL folder,
to your dictionnaire if you have one,
to where you want the results to be saved
and to your meddra files.

### From sample

```{r path_classic, eval = FALSE}
# #### IMPORT #### ####

# ---- Paths sample ---- ####
path_base   <- "~/Database/sample/"
path_who    <- "~/Database/who/"

path_dic    <- "~/my_project/generic_dictionnaire.R"
# alternative with the here syntax
path_dic    <- here("generic_dicitionnaire.R")

path_res    <- "~/my_project/my_res/"
# alternative
path_res    <- here("my_res")

path_meddra <- "~/Database/meddra_25_1/meddra_25_1/english/MedAscii/"
```
### From serveur files

If you're using the Caen CHU server, you paths might look like this

```{r path_server, eval = FALSE}
path_base   <- "/work/pharma/vigibase_2021_sep/main/"
path_who    <- "/work/pharma/vigibase_2021_sep/who/"
path_dic    <- "/home/dolladille-c/sclerodermia/03_dictionnaire.R"
path_res    <- "/home/dolladille-c/sclerodermia/"
path_meddra <- "/home/dolladille-c/meddra_24_0/english/medascii/"

```

### Load datasets

```{r classic_datasets, eval = FALSE}
demo     <- dt_fst(path_base, "demo")
adr      <- dt_fst(path_base, "adr")
drug     <- dt_fst(path_base, "drug")
link     <- dt_fst(path_base, "link")
out      <- dt_fst(path_base, "out")
srce     <- dt_fst(path_base, "srce")
followup <- dt_fst(path_base, "srce")

suspdup  <- dt_fst(path_base, "suspectedduplicates")

thg <- dt_fst(path_who, "THG")
mp_short <- dt_fst(path_who, "mp_short")
```

### Work with preloaded test tables

Si vous n'avez pas de base pre-construite, servez vous des tables prechargees. 
(cf infra) Dans cette vignette, c'est ce que nous allons faire.

```{r test_datasets}
demo     <- demo_
adr      <- adr_
drug     <- drug_
link     <- link_
out      <- out_
followup <- followup_

srce     <- srce_

thg      <- thg_
mp_short <- mp_short_
meddra   <- meddra_
smq_list_content <- smq_list_content_

suspdup <- 
  data.table(
    UMCReportId = 1,
    SuspectedduplicateReportId = NA
  )
```

## Dictionary

> !! Wait a minute!

Now you will "source" your dictionnaire...

Do you have one? There's a dedicated vignette here 
\code{vignette("generic_dictionary", package = "pharmacocaen")}.

If its the first time you write this script, you may want
to **STOP** here and create one.

That's even most interesting as this vignette will use items created from the
dictionary, and you may get lost if you haven't seen them before.

Once you're done in generic_dictionnaire.R, move back here.

```{r dict, eval=FALSE}
source(path_dic)
```

### Don't have a dictionary yet or working with test sets?

```{r codes, eval=TRUE}
d_drecno <- ex_$d_drecno

atc_drecno <-
  get_atc_code(atc_sel =  rlang::list2(l03_j01 = c("L03AA", "J01CA")),
               mp_short = mp_short_,
               thg_data = thg_,
               vigilyze = TRUE)

a_llt <- ex_$a_llt
```

## DEMO data management

```{r demo_dm}
# ---- Deduplicating ---- ####

demo <- demo[!(UMCReportId %in% suspdup[, SuspectedduplicateReportId]),]

# ---- Drugs ---- ####

# From a list of drugs

demo <-
  demo %>%
  add_drug(
    d_code = d_drecno,
    drug_data = drug
  )

# From ATC

demo <-
  demo %>%
  add_drug(
    d_code = atc_drecno,
    drug_data = drug
  )

# ---- Reactions ---- ####

demo <-
  demo %>%
  add_adr(
    a_code = a_llt,
    adr_data = adr
  )

# ---- Demographics ---- ####

# Age, sex

demo <-
  demo %>%
  mutate(
    age = cut(as.integer(AgeGroup),
              breaks = c(0,4,5,6,7,8),
              include.lowest = TRUE, right = TRUE,
              labels = c("<18", "18-45","45-64", "65-74", "75+")),

    sex = case_when(Gender == "1" ~ 1,
                    Gender == "2" ~ 2,
                    Gender %in% c("-","0","9") ~ NA_real_,
                    TRUE ~ NA_real_)
  )

# Death + disponibilite de l'outcome

demo <-
  demo %>%
  mutate(death =
           if_else(UMCReportId %in% out[, UMCReportId],
                   UMCReportId %in% out[Seriousness == "1", UMCReportId],
                   NA)
  )

# follow-up, seriousness

demo <-
  demo %>%
  mutate(
    fup = if_else(UMCReportId %in% followup[,UMCReportId], 1, 0),
    serious = if_else(UMCReportId %in% out[, UMCReportId],
                     UMCReportId %in% out[Serious == "Y", UMCReportId],
                     NA)
  )

# year
demo[, `:=` (year = as.numeric(substr(FirstDateDatabase, start = 1, stop = 4)))]

# type of reporter
demo %<>% left_join(srce[, .(UMCReportId, type_reporter = Type)], by = "UMCReportId") %>% data.table
```

### Check your data management in demo

```{r demo_check, eval = FALSE}
demo %>%
  check_dm(cols = c(names(d_drecno), names(a_llt), "fup"))
```

## LUDA 

### Create LUDA

Create the LUDA table (a link table joined with UMCReportId). For convenience, it is also best practice to create tto_mean and range variables at this stage (at least in my opinion).

```{r luda_create}
luda <-
  link %>% 
  left_join(
    drug %>% 
      select(UMCReportId, Drug_Id),
    by = "Drug_Id"
  ) %>%
    mutate(
    tto_mean = (TimeToOnsetMax + TimeToOnsetMin) / 2,
    range = (TimeToOnsetMax + TimeToOnsetMin) / 2 - TimeToOnsetMin
  ) 
```

### Data managed LUDA

```{r luda_drug_adr}
luda <-
  luda %>% 
    add_drug(
    d_code = d_drecno,
    drug_data = drug,
    data_type = "link"
    ) %>% 
    add_adr(
    a_code = a_llt,
    adr_data = adr,
    data_type = "link"
    )
```

## Basic models

```{r mods}
# ---- Bivariate ---- ####

rb <-
  purrr::map_dfr(
    names(d_drecno),
    function(x)
      purrr::map_dfr(
        names(a_llt),
        function(y)
          demo %>%
          compute_or_abcd(x = x,
                          y = y)
        )
  )

# remove the hashes to save your results
## write.csv2(rb, paste0(path_res, "rb.csv"), row.names = FALSE)

# ---- Multivariate ---- ####

mod <-
  glm(a_colitis ~ nivolumab + age + sex,
      family = "binomial",
      data = demo)

rm <-
  summary(mod)$coefficients %>%
  compute_or_mod(
    estimate = Estimate,
    std_er = Std..Error
    )

# remove the hashes to save your results
## write.csv2(rm, paste0(path_res, "rm.csv"), row.names = FALSE)
```

## Basic descriptive

```{r desc_demo}
r_desc <- 
  demo %>% 
  desc_facvar(
    vf = c("age", "sex",
           "type_reporter",
           "Type",
           "year",
           names(d_drecno),
           names(a_llt),
           "serious", "death"),
    ncat_max = 20
  )
```

```{r desc_tto}
r_tto <-
  desc_tto(
    luda,
    adr_s = names(a_llt),
    drug_s = names(d_drecno)
    ) 
```

```{r desc_dch}
r_dch <-
  desc_dch(
    luda,
    adr_s = names(a_llt),
    drug_s = names(d_drecno)
    )
```

```{r desc_rch}
r_rch <-
  desc_rch(
    luda,
    demo_data = demo,
    adr_s = names(a_llt),
    drug_s = names(d_drecno)
  )
```
