---
title: "Getting started"
output: 
  rmarkdown::html_vignette:
    keep_md: true
    toc: true
vignette: >
  %\VignetteIndexEntry{getting_started}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(vigicaen)
```

# Purpose

Have you subscribed to VigiBase Extract Case Level, WHODrug and MedDRA?

Congratulations! You've made it half way to explore VigiBase.

Now, you need to process those large text/ascii files into R readable
files. This is the purpose of this vignette.

## Overview

 The `tb_*` functions (Table Builders) will help you through the process.

| `tb_*` function | Usage                                                                    |
|--------------------|----------------------------------------------------|
| `tb_vigibase`     | Main tables (demo, drugs, adverse drug reactions), and subsidiary tables |
| `tb_who`          | WHODrug tables                                                           |
| `tb_meddra`       | MedDRA tables                                                            |

# Deploy tables

## Folder and files

First, UNZIP the files you received from UMC (beware that you also need
to unzip **sub-folders**, inside the main zip file).

That will give you a folder structure like this:

-   /vigibase_ecl
-   .../main
-   .../who
-   .../sub

Gather your meddra files into a single folder, ideally close to the
vigibase_ecl one.

-   /meddra

Although the names can be longer (e.g. "main_sep_1_24"), 
we **strongly** suggest you shorten them
to `main`, `who`, `sub` for Vigibase, and `meddra` for MedDRA.

## Table builders (`tb_*` functions)

Provide a path to these files on your computer.

We usually create `path_*` objects in R, then use them in functions.
This is optional, you can also pass path directly to `tb_*` functions.


```{r basic, eval = FALSE}

# change for the correct paths on your computer
path_base   <- "../vigibase_ecl/main" 
path_sub    <- "../vigibase_ecl/sub"
path_who    <- "../vigibase_ecl/who"
path_meddra <- "../meddra"

tb_vigibase(
  path_base = path_base, 
  path_sub  = path_sub
  )

tb_who(
  path_who  = path_who
  )

tb_meddra(
  path_meddra = path_meddra
  )
```

> Deploying table have to be done ONCE per database version.

New parquet files will be created in the same directories as provided
to `path_*`.

> The suspectedduplicates table is delivered with subsidiary files, but
> the `tb_vigibase()` function will output it in the main directory,
> under a slightly modified name.

## Outputs

The `tb_*` functions create
[parquet](https://arrow.apache.org/docs/r/index.html) files. Parquet is
an open-source format, supported by Arrow.

Not all tables from each source are created during the process.

-   Main: All tables + subsidiary tables (from sub):
demo, drug, adr, link, ind, srce, out, followup, suspdup,
AgeGroup, Dechallenge, Dechallenge2, Frequency,
Gender, Notifier, Outcome, Rechallenge, Rechallenge2, Region,
RepBasis, ReportType, RouteOfAdm, Seriousness, and SizeUnit.

-   WHODrug: All tables: mp, thg, pp, 
ing, srce, org, ccode, atc, sun, pf, str, prg, prt, unitx. Again,
check correspondence in the delivered files from UMC.

-   Meddra: meddra_hierarchy, smq_list, and smq_content.

`mp` is modified in comparison to the original file, drug names were lowered case and
trimed, to be more handy. A few columns have also been
removed.

# Loading tables into R

You can use the arrow package built-in readers to load the tables, or
the `dt_parquet()` function from this package, which is a shorthand for
`data.table::as.data.table(read_parquet())`.

```{r load_tables, eval = FALSE}

demo   <- dt_parquet(path_main, "demo")
drug   <- dt_parquet(path_main, "drug")
adr    <- dt_parquet(path_main, "adr")
link   <- dt_parquet(path_main, "link")
# etc.

mp     <- dt_parquet(path_who, "mp")

meddra <- dt_parquet(path_meddra, "meddra_hierarchy")

```

> Parquet files will be loaded IN memory, by default. To change this
> behavior, remember to set `in_memory = FALSE`. Be careful that the
> output format will be different.

# Working on a computer with low specifications

The main ressource to store objects in R is the RAM. If your computer
don't have much RAM (e.g., only 16GB), you will not be able to load
all Vigibase tables at once in your session.

In this case, you can use the `in_memory` argument of `dt_parquet()` to
load tables **OUT** of memory.

```{r load_tables_out_memory, eval = FALSE}
demo   <- dt_parquet(path_main, "demo", in_memory = FALSE)
```

This will allow you to go through most of the data management steps.

At the time of statistical modelling, you might need to load a table
and/or use modelling packages that do not rely on RAM (one example:
[biglm](https://cran.r-project.org/web/packages/biglm/index.html))

# Subsetting tables

If you want to work on a specific subset of VigiBase (e.g. older adults,
some specific drugs, or adverse drug reaction), you can create a
dedicated subset with the `tb_subset()` function.

Currently, you can subset on

-   Age groups

-   Drug record numbers (DrecNo) and Medicinal product IDs
    (MedicinalProd_Id)

-   MedDRA codes of adverse drug reactions (MedDRA_Id)

Visit `tb_subset()` help page for more information.

You will need to collect drug or reaction **IDs** if you want to subset
on these. See `vignette("basic_workflow")`.

```{r tb_subset, eval = FALSE}

sv_selection <-
    c(7, #  65 - 74 years group
      8) #  >= 75 years group

wd_in <- "some/place/on/your/computer/containing/vigibase_ecl/main"

wd_out <- paste0(wd_in, "/", "more_than_65_subset", "/")

tb_subset(wd_in, wd_out,
          subset_var = "age",
          sv_selection = sv_selection)
```
