---
title: "Routine pharmacovigilance"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{routine_pharmacovigilance}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(vigicaen)
```

# Motivation

The aim of this vignette is to describe how to use vigicaen in the
context of routine pharmacovigilance.

In a hurry? Check the condensed script of this vignette at
`vignette("routine_script")`.

The standard user of this vignette is:

-    A pharmacovigilance practitioner
-    Interested in R, but not necessarily willing to become an R expert,
nor a data management expert.
-    Working on a pharmacovigilance case, and seeking for additional data, to
improve his/her case report: Disproportionality Estimates, possibly refined
to a certain context, and time from drug initiation to reaction onset.

Meeting these criteria? Then this vignette is for you!

# The whole game

The idea is to reduce to the minimum steps to get data from VigiBase.

1.    Load VigiBase data into R

2.    Identify the drug and reaction of interest

3.    Run `vigi_routine()`

4.    Save the results, in order to insert them into your pharmacovigilance
report.

`vigi_routine()` is intended as the cornerstone function in this workflow.

> We will use the following example: Say a physician reported a case of 
> colitis with nivolumab. We would like to know if there is a reporting
> association between colitis and nivolumab, and we would like to 
> analyze time to onset.

Obviously, the first part is more interesting if the association between
colitis and nivolumab is not well established.

# Load tables

If you haven't already, you should build tables in an R compatible format,
using the `tb_vigibase()` and `tb_who()` functions. More information
can be found in  `vignette("getting_started")`.

This rather long step must only be done once per database version.

Then, you need to set up the paths to your tables, and use
`dt_parquet()`.

```{r paths}
path_base   <- "~/vigibase/main/"
path_who    <- "~/vigibase/who/"
path_meddra <- "~/meddra/"
```

Load tables into R. Preferably, let them out of memory, especially
if you have a computer with rather low specifications (e.g., 16GB of RAM).

```{r load_tables, eval=FALSE}
demo <- dt_parquet(path_base, "demo", in_memory = FALSE)
drug <- dt_parquet(path_base, "drug", in_memory = FALSE)
adr  <- dt_parquet(path_base, "adr", in_memory = FALSE)
link <- dt_parquet(path_base, "link", in_memory = FALSE)

mp <- dt_parquet(path_who,  "mp")
meddra   <- dt_parquet(path_meddra, "meddra")
```

For this vignette, we will use built-in example tables instead.

```{r load_example_tables}
demo     <- demo_
adr      <- adr_
drug     <- drug_
link     <- link_

mp <- mp_
meddra   <- meddra_
```

# Identify drug and reaction of interest

## Select the drug and reaction

First, create named lists of the drug and reaction of interest. This is 
exactly similar to the `vignette("basic_workflow")`.

```{r a_sel_d_sel}

d_sel <- list(
  nivolumab = "nivolumab"
)

a_sel <- list(
  # this is a High Level Term
  colitis = "Colitis (excl infective)"
)

```

When working with vigicaen, we always need to tell R which drug and which
reaction we are interested in. Here, we create the `d_sel` object, which is a 
list, with one item. The item name is nivolumab, and the content of this item
is "nivolumab". The same goes for `a_sel`. Mind the capital letter at the 
beginning of "**C**olitis": It must match exactly to the MedDRA term of this
reaction.

The `d_sel` must contains an "active ingredient", which is the same as the
international non-proprietary name, in a majority of cases.

The `a_sel` must contain a MedDRA term, that can be either an Preferred Term,
or any other Term level in MedDRA. It could also be an SMQ.

> In order to use `vigi_routine()`, both `d_sel` and `a_sel` must
> contain one, and only one, item each.

## Collect IDs of drug and reaction

The `get_*` functions family let you convert these terms into IDs, that
can be used in VigiBase tables.

For drugs, we need to get DrecNos (Drug Record Numbers). For reactions, we need
low-level term codes.

In this example, we will use `get_drecno()` and `get_llt_soc()`.

Feel free to browse the other `get_*` functions: `get_llt_smq()`, 
`get_atc_code()`.

```{r get_drecno_get_llt_soc}
d_code <- 
  get_drecno(d_sel, mp = mp)

a_code <-
  get_llt_soc(a_sel, term_level = "hlt", meddra = meddra)
```

# Use `vigi_routine()`

You're almost done, the last thing you need to do is run `vigi_routine()`.

It takes several arguments:

-    `demo_data`, `drug_data`, `adr_data`, `link_data`: 
The tables you loaded earlier (as demo, drug, adr, and link)

-    `d_code`, `a_code`: The IDs of the drug and reaction of interest

-    `vigibase_version`: It's a character string that will be printed
in the graph legend. You can use anything, such as "September 2024".
Just remember to write _something_.

```{r vigi_routine, fig.height=6.7, fig.width=4}
#| fig.alt: >
#|   Example of vigi_routine.

vigi_routine(
  demo_data = demo,
  drug_data = drug,
  adr_data  = adr,
  link_data = link,
  d_code    = d_code,
  a_code    = a_code,
  vigibase_version = "September 2024"
)
```

Lets take a look at the results. There are 2 graphs, one of top of the other.

-    The first graph is a disproportionality analysis, using the 
Information Component. The color scale is purely indicative and should
not be considered as standard.

-    The second graph is a time to onset analysis. It shows the distribution
of time to onset from drug initiation (here, nivolumab), to reaction onset
(colitis). The scale is log-transformed, and span from 1 day to ... 10 years!
The pale blue and sustained blue bars at the bottom indicate the number of 
limits where 50% and 80% of the patients fall. We can roughly see that 50% of
the patients had a time to onset comprised between ~1.5 month and 1 year.
This second graph uses only cases where the drug was **suspected** to be responsible
for the reaction.

## Add your own case time to onset

As we said earlier, we are working on a case reported by a physician. In
this case, the colitis happened 150days after nivolumab initiation.

We can add this information to the graph, using the `case_tto` argument.


```{r vigi_routine_case, fig.height=6.7, fig.width=4}
#| fig.alt: >
#|   Example of vigi_routine with case data.

vigi_routine(
  case_tto  = 150,
  demo_data = demo,
  drug_data = drug,
  adr_data  = adr,
  link_data = link,
  d_code    = d_code,
  a_code    = a_code,
  vigibase_version = "September 2024"
)
```

## Customize the graph

There are a few options to customize the output of this graph.

You can use the `d_label` and `a_label` arguments to change the display in
the legend.

## Exporting your results

You can save the results of `vigi_routine()` in a separate file, using
the `export_to` argument.

`export_to` must be a character string, indicating the path on your computer
where you want to file to be saved. It **must** end with one of the following
extensions:

> ".eps", ".ps", ".tex" (pictex), ".pdf", ".jpeg",
> ".tiff", ".png", ".bmp", ".svg" or ".wmf" (windows only)

Some of these extensions (like .svg) require additional packages to work.
(e.g. svglite for ".svg").

```{r vigi_routine_case_export, eval=FALSE}
vigi_routine(
  case_tto  = 150,
  demo_data = demo,
  drug_data = drug,
  adr_data  = adr,
  link_data = link,
  d_code    = d_code,
  a_code    = a_code,
  d_label   = "Nivolumab",
  a_label   = "Colitis (HLT)",
  vigibase_version = "September 2024",
  export_to = "~/vigicaen_graph.png"
)
```
