---
title: "Dictionary"
output: 
  rmarkdown::html_vignette:
    keep_md: true
    toc: true
vignette: >
  %\VignetteIndexEntry{template_dictionary}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}

---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Dictionary script template

You may want to organize your research folder using small scripts,
each designed for a part of your analysis plan.

Inspired from [Reproducible Research in R](https://mjrolland.github.io/reproducibility-checklist/cheatsheet.html)

-    /folder
-    ../data
-    ../scripts
-    ../etc.

The dictionary might be located in scripts, under the name "dict.R".

-    ../scripts/00_dict.R

If you haven't, visit `vignette("getting_started")` and `vignette("basic_workflow")` for details on what this template does.

```{r libraries, warning = FALSE, message = FALSE}
library(vigicaen)
library(rlang)
library(dplyr)
```

```{r real_data, eval = FALSE}
# #### PATHS #### ####

path_who    <- "../vigibase_ecl/who" # adapt to your own

path_meddra <- "../meddra" 

# #### IMPORT #### ####

# ---- Datasets ---- ####

mp     <- dt_parquet(path_who, "mp")

thg    <- dt_parquet(path_who, "thg")

meddra <- dt_parquet(path_meddra, "meddra_hierarchy")

smq_list <- dt_parquet(path_meddra, "meddra_smq_list")

smq_content <- dt_parquet(path_meddra, "meddra_smq_content")
```

```{r test_data, echo = FALSE, include = FALSE}
meddra <- meddra_
mp <- mp_
smq_list <- smq_list_
smq_content <- smq_content_
thg <- thg_
```

```{r drugs_adrs, echo = TRUE, eval = TRUE, message = FALSE}
# #### Drug definitions #### ####

d_sel <-
  list(
    nivolumab = "nivolumab", # use lower case
    ici = c("nivolumab", "pembrolizumab")
  )

# ---- ID collector

d_drecno <-
  d_sel |>
  get_drecno(
    mp = mp
  )

# #### ATC definitions #### ####

atc_sel <-
  list(l03 = "L03") # ATC code

# ---- ID collector

atc_drecno <-
  atc_sel |>
  get_atc_code(
    mp = mp,
    thg_data = thg
  )

# #### ADR definitions #### ####

# adr selection, matching meddra case

# ---- SOC definition

pt_sel <-
  list(
    hepatitis = c("Hepatitis", "Budd-Chiari syndrome"),
    hyponatremia = c("Pneumonitis")
  )

hlt_sel <-
  list(
    conduction = "Diarrhoea (excl infective)"
  )

# ---- SMQ definition

smq_sel <-
  list(
    hta = "Ischaemic heart disease (SMQ)"
  )

# ---- ID collectors

a_llt <-
  c(
    get_llt_soc(pt_sel,
                term_level = "pt",
                meddra = meddra),
    get_llt_soc(hlt_sel,
                term_level = "hlt",
                meddra),
    get_llt_smq(smq_sel,
                smq_scope =  "narrow",
                smq_list,
                smq_content)
  )
```

